---
title: "FredCast"
author: "Ramon Prat"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
---
# My Monthly Forecast of USA Macroeconomic Indicators
```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(timeSeries)
library(forecast)
library(jsonlite)
library(quantmod)
library(forecast)
library(xtable)
```


```{r download, include=FALSE}
#json config to download data from fred
#api_key<-"&api_key=9912ee381321a596b999f26cfa5da2a5"
setwd("~/Documentos/Madimon/Investment/econ_outlook/fcasts/reports")
api_key<-source("fed_api.R")
api_key<-api_key$value
url<-"https://api.stlouisfed.org/fred/series/observations?"
file_type<-"&file_type=json"
```

```{r employment, cache=FALSE,include=FALSE}
#Change in nonfarm payroll employment
s1<-"series_id=PAYEMS" #seasonally adjusted monthly in nonfarm payroll employment to the nearest thousand persons
s1<-fromJSON(paste0(url, s1, api_key,file_type))
s1_df<-as.data.frame(s1)
charvec<-as.Date(as.character(s1_df[,15]))
data<-as.numeric(s1_df[,16])
s1<-as.timeSeries(data, charvec)
s1<-diff(s1,1) #to compute monthly change
date_s1<-last(s1_df$observations.date)
value_s1<-last(s1_df$observations.value)
date_s1
tail(s1)
```


```{r inflation, cache=FALSE,include=FALSE}
s2<-"series_id=CPIAUCSL"    #"seasonally adjusted annualized monthly percentage change in headline CPI
                            #rounded to the nearest tenth of a percent"Inflation is usually reported in terms of year to year changes. In this game, you will be considering monthly changes at an annualized rate. Often, the monthly change in the CPI reflects short-term price changes in goods and services with volatile prices.
s2<-fromJSON(paste0(url, s2, api_key,file_type))
s2_df<-as.data.frame(s2)
charvec<-as.Date(as.character(s2_df[,15]))
data<-as.numeric(s2_df[,16])
s2<-as.timeSeries(data, charvec)
#tail(s2)
s2<-((s2/lag(s2,1))^12-1)*100 #annualized percentage change 
date_s2<-last(s2_df$observations.date)
tail(s2)
date_s2
```


```{r GDP, cache=FALSE,include=FALSE}
#Real GDP Growth
s3<-"series_id=A191RL1Q225SBEA"     #"seasonally adjusted annualized quarter-to-
                                    #quarter percentage change advance estimate 
                                    #of GDP from the previous final estimate of 
                                    #GDP, rounded to the nearest tenth of a percent"
s3<-fromJSON(paste0(url, s3, api_key,file_type))
s3_df<-as.data.frame(s3)
charvec<-as.Date(as.character(s3_df[,15]))
data<-as.numeric(s3_df[,16])
s3<-as.timeSeries(data, charvec)
date_s3<-last(s3_df$observations.date)
tail(s3)
#date_s3
#
```

```{r unemployment, include=FALSE,cache=FALSE}
#"Unemployment 
s4<-"series_id=UNRATE"      #is the seasonally adjusted level of the headline
                            #unemployment rate in percentage points rounded to 
                            #the nearest tenth of a percent")
s4<-fromJSON(paste0(url, s4, api_key,file_type))
s4_df<-as.data.frame(s4)
charvec<-as.Date(as.character(s4_df[,15]))
data<-as.numeric(s4_df[,16])
s4<-as.timeSeries(data, charvec)
date_s4<-last(s4_df$observations.date)
#tail(s4)
#date_s4
```


```{r civilian_labor_force_participation, include=FALSE,cache=FALSE}
#Participation 
s5<-"series_id=CIVPART"      #is the seasonally adjusted level Percent, Seasonally Adjusted
s5<-fromJSON(paste0(url, s5, api_key,file_type))
s5_df<-as.data.frame(s5)
charvec<-as.Date(as.character(s5_df[,15]))
data<-as.numeric(s5_df[,16])
s5<-as.timeSeries(data, charvec)
date_s5<-last(s5_df$observations.date)
tail(s5)
date_s5
```




```{r output, include=FALSE, cache=FALSE}
# FORECAST

fit1 <- auto.arima(s1)
fit2 <- auto.arima(s2)
fit3 <- auto.arima(s3)
fit4 <- auto.arima(s4)
fit5 <- auto.arima(s5)

ts1_cast3<-forecast(fit1, 3)
ts2_cast3<-forecast(fit2, 3)
ts3_cast3<-forecast(fit3, 3)
ts4_cast3<-forecast(fit4, 3)
ts5_cast3<-forecast(fit5, 3)

acc1<-accuracy(fit1)
acc2<-accuracy(fit2)
acc3<-accuracy(fit3)
acc4<-accuracy(fit4)
acc5<-accuracy(fit5)
me1<-acc1[1,1]
me2<-acc2[1,1]
me3<-acc3[1,1]
me4<-acc4[1,1]
me5<-acc5[1,1]
me<-round(c(me1,me2,me3,me4,me5),4)

parameter<-c("Employment", "Inflation", "RealGDP", "Unemployment","LaborParticipation")
avdate<-c(date_s1,date_s2,date_s3,date_s4, date_s5)
value3<-round(c(ts1_cast3$mean,ts2_cast3$mean,ts3_cast3$mean,ts4_cast3$mean, ts5_cast3$mean),4)
value3<-as.data.frame(matrix(c(value3),nrow=5,ncol=3,byrow=TRUE))
last<-round(c(last(s1),last(s2),last(s3), last(s4), last(s5)),4)
freq<-c("monthly","monthly","quarterly","monthly", "monthly")
forecast<-data.frame(parameter,freq, last,avdate, value3, me)#%, unit)
colnames(forecast) <- c("Parameter", "freq","last","AvDate", "Fcast+1","Fcast+2","Fcast+3","ME")#, "Desc")
```

```{r report, results="asis", echo=FALSE}
xt<-xtable(forecast, comment=FALSE)
print(xt, type="latex", comment=FALSE)
```
[FredCast](https://research.stlouisfed.org/useraccount/fredcast/forecast "FredCast")

# Series

## Payroll Employment
Monthly Change, Seasonally Adjusted (rounded to nearest thousand persons)
Payroll Employment is an estimate of the number of workers U.S. firms employ and send paychecks to every month. Total Nonfarm, commonly known as Total Nonfarm Payroll, is a measure of the number of U.S. workers in the economy that excludes proprietors, private household employees, unpaid volunteers, farm employees, and the unincorporated self-employed. This measure accounts for approximately 80 percent of the workers who contribute to Gross Domestic Product (GDP). The change in payroll employment is the number of jobs added or lost over a month's time. On `r last(s1_df$observations.date)` Payroll Employment was `r last(s1_df$observations.value)` thousands of people, a `r last(s1)` change from `r s1_df$observations.date[length(s1_df$observations.date)-1]`.



```{r, echo=FALSE, message=FALSE}
plot(forecast(fit1,12),main="Payroll Employment Monthly Change",36,ylab="thousands",ylim=c(-100,+400),cex.lab=0.75)
```

## Civil Force Participation
Measures the labor force as a percentage of the civilian noninstitutional population, that is people 16 years of age and older residing in the 50 states and District of Columbia, who are not inmates of institutions (e.g., penal and mental facilities, homes for the aged) and who are not on active duty in the Armed Forces.

```{r, echo=FALSE, message=FALSE}
plot(forecast(fit5,24),main="Civil Force Participation",436,ylab="%",cex.lab=0.75)
abline( h = mean(as.numeric(s5_df$observations.value)))
text(paste("Mean =", round(mean(as.numeric(s5_df$observations.value)), 1), "\nMedian =", round(median(as.numeric(s5_df$observations.value)), 1), "\nStd.Dev =", round(sd(as.numeric(s5_df$observations.value)))))
```

## Real GDP Growth
Quarterly, Percent Change from Preceding Period, Seasonally Adjusted Annual Rate (rounded to nearest .1%).
Real Gross Domestic Product (GDP) is the inflation-adjusted value of final goods and services produced by an economy over a period of time. Real GDP growth is the percentage change in GDP from one period to the next.
GDP is reported as if the value of final goods and services produced in a quarter had continued over the entire year. This process, called annualization, means the quarterly growth rate for real GDP is actually the percent change between two annualized numbers, producing an annual rate.
```{r, echo=FALSE, message=FALSE}
plot(forecast(fit3,12),main="GDP",36,ylab="%",ylim=c(-5,5),cex.lab=0.75)
```


## Unemployment
Monthly, Seasonally Adjusted (rounded to the nearest 0.1%). Unemployment caused by changes in the weather or seasons is removed by government agencies to make the underlying trend in the data easier to observe.
```{r, echo=FALSE, message=FALSE}
plot(forecast(fit4,12),main="Unemployment",36,ylab="%",ylim=c(4.5,7.5),cex.lab=0.75)
```


## CPI Inflation
Monthly Compounded Annual Rate of Change, Seasonally Adjusted (rounded to nearest 0.1%).
The Consumer Price Index (CPI) is a measure of the level of prices of a collection of commonly purchased goods and services (e.g., housing, transportation, food and medical care) in an economy over time. The percentage change in the CPI from one period to the next measures inflation.
Inflation is usually reported in terms of year to year changes. In this game, you will be considering monthly changes at an annualized rate. Often, the monthly change in the CPI reflects short-term price changes in goods and services with volatile prices.
```{r, echo=FALSE, message=FALSE}
plot(forecast(fit2,12),main="Inflation",36,ylab="%",ylim=c(-4,4),cex.lab=0.75)
```

